---
title: "HW 2 Student"
author: "Andy Ackerman"
date: "10/17/2023"
output: 
  html_document:
    number_sections: true
---

This homework is meant to illustrate the methods of classification algorithms as well as their potential pitfalls.  In class, we demonstrated K-Nearest-Neighbors using the `iris` dataset.  Today I will give you a different subset of this same data, and you will train a KNN classifier.  

```{r, echo = FALSE}
set.seed(123)
library(class)

df <- data(iris) 

normal <-function(x) {
  (x -min(x))/(max(x)-min(x))   
}

iris_norm <- as.data.frame(lapply(iris[,c(1,2,3,4)], normal))

subset <- c(1:45, 58, 60:70, 82, 94, 110:150)
iris_train <- iris_norm[subset,] 
iris_test <- iris_norm[-subset,] 

iris_target_category <- iris[subset,5]
iris_test_category <- iris[-subset,5]


```

#
Above, I have given you a training-testing partition.  Train the KNN with $K = 5$ on the training data and use this to classify the 50 test observations.  Once you have classified the test observations, create a contingency table -- like we did in class -- to evaluate which observations your algorithm is misclassifying.   

```{r}
set.seed(123)
#STUDENT INPUT
pr <- knn(train = iris_train, test = iris_test, cl = iris_target_category, k = 5)
tab <- table(pr, iris_test_category)
tab

accuracy <- function(x){
  sum(diag(x))/sum(rowSums(x))*100
}

accuracy(tab)
```

#

Discuss your results.  If you have done this correctly, you should have a classification error rate that is roughly 20% higher than what we observed in class.  Why is this the case? In particular run a summary of the `iris_test_category` as well as `iris_target_category` and discuss how this plays a role in your answer.  

*STUDENT INPUT* 
In my above code, I find that there are 11 more virginicas than in actuality. This results in a misclassification rate of 78%, and this is largely the case due to the lack of random sampling for the training/test partition. Below, I run a summary of `iris_test_category` and `iris_target_category`.

```{r}
summary(iris_test_category)
summary(iris_target_category)
```

From this summary, it is clear that the target category has a vastly different distribution of the three species of flowers than the testing data, resulting in a poor ability to distinguish the three when dealing with test data. Had these been more representative, it is likely that knn would be more able to differentiate the three species. 

#

Build a github repository to store your homework assignments.  Share the link in this file.  

*STUDENT INPUT*

https://github.com/samplle/STOR-390-HW/tree/main

